---
title: "Bee Colony Loss Data"
format: html
editor: visual
---

```{r}

library(tidyverse)
library(tidymodels)
library(skimr)




colony <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-01-11/colony.csv')

```

```{r}
colony_splits <- initial_split(colony, prop = 0.5)

training_data <- training(colony_splits)
test_data <- testing(colony_splits)
```

# Bee Colony Loss Analysis

## Abstract

This is an analysis of the Bee Colony data set. The bee colony data set provides information on honey bee colonies across the United States with regards to the number of colonies, maximum, lost, percent lost, added, renovated, and percent renovated. In this technical report I aim to answer the following questions...... This data set spans several years during which the number of bee colony losses per state were observed and recorded. Using RStudio we were able to create box plots, allowing for better visualization and analysis of the data when focusing on correlation between months/year or state and bee colony losses observed. Consider adding answers to your questions...

## Interesting Questions

-   During which months did we lose the most bee colonies? The least?

-   During which years did we lose the most bee colonies? The least?

-   Is there a state that seems superior to the others in terms of colony size?

-   Is there a state that seems to lose more bee colonies than others?

## Hypotheses

-   I hypothesize that the most bee colonies were lost throughout October-December and the least bee colonies were lost throughout April-June.

-   I hypothesize that the most bee colonies were lost during 2015 and the least bee colonies were lost during 2017.

-   I hypothesize that California is the largest in size and has the most bee colonies.

-   I hypothesize that Florida lost the most bee colonies during the years this data was collected.

```{r}
training_data %>%
  skim()
```

## Introduction

This report provides information on honey bee colonies in the US regarding the number of colonies, maximum, lost, percent lost, added, renovated, and percent renovated. Colony loss rates are calculated as the ratio of the number of colonies lost to the number of colonies managed over a defined period. Colony loss rates are best interpreted as a turn-over rate, as high levels of losses do not necessarily result in a decrease in the total number of colonies managed in the United States. I hypothesized that the most bee colonies were lost October-December and the least number of bee colonies were lost April-June. It was also hypothesized that the most bee colonies were lost during 2015 and the least bee colonies were lost during 2017, when focusing on year.

## Exploratory Data

Format:

-   Here's what I am about to do...

-   Here's why...

-   Here's the thing I did (the picture)

-   Here's what we are seeing (or not seeing)

-   Here's what I'll do next...repeat...

...Moved table here...

```{r}
training_data %>%
  filter(state == "Florida") %>%
  skim()
```

This table allows me to look at data specific to the state of Florida. This table tells me that Florida has a mean value of 13% for percent of average colony lost. This is a slight increase from the first table created at the beginning of the report, which gave a value of 11% for the average percent of honey bee colonies lost. Seeing how both values are precise makes me more confident of this data.

...Original start...

```{r}

training_data %>%
  ggplot() +
  geom_bar(aes(x=months, y= colony_lost),
           stat="identity")
  

```

This graph made me realize that a bar graph was not the best format to use when displaying my data. Although the data could be organized in a better way, this graph told me that the most bee colonies were lost: April-June, January-March, July-September, and October-December. The most bee colonies were lost October-December.

```{r}

training_data %>%
  ggplot() +
  geom_boxplot(aes(x=months, y= colony_lost)) +
  scale_y_log10()
```

This graph, especially when the scale_y\_log10() was added to the code, displayed the data in a more visually appealing way. Not only were we able to look at the box-plot and see that the most bee colonies were lost October-December, but we also learned that the amount of bee colonies lost were very numerically close between the different months of time. This makes me a little wary of the data because it makes me question how accurate the data is, I would have assumed there to be a more drastic difference between the months (more specifically Winter and Spring/Fall/Summer).

```{r}
training_data %>%
  ggplot() +
  geom_boxplot(aes(x=colony_lost, y= year)) 
  
```

This box-plot was my first attempt at looking at the (possible) correlation between amount of bee colonies lost and the year. I was curious to see if there was a year were honey bees were more greatly affected. Unfortunately, this graph is only recognizing the year 2018.

```{r}
training_data %>%
  ggplot() +
  geom_point(aes(x=colony_lost, y=colony_added)) +
  geom_smooth(aes(x=colony_lost, y=colony_added),
              method="lm", se= FALSE, color="red")

```

This graph shows the correlation between the amount of honey bee colonies lost and amount of honey bees added. A line of best fit was added to better analyze and understand the data. Looking at the line of best fit we see that the lower numbers appear more accurate compared to higher data sets where there are more outliers present.



## Conclusion
